% Auto-generated from tab_EA20_price_counterfactual_measures_A2_like.csv.
\begin{landscape}
\scriptsize
\setlength{\LTleft}{0pt}
\setlength{\LTright}{0pt}
\setlength{\tabcolsep}{2pt}
\renewcommand{\arraystretch}{1.12}
\begin{longtable}{@{}p{0.55in}p{1.65in}p{1.05in}p{3.35in}p{1.25in}@{}}
\caption{EA20 price measures and simulated counterfactuals}\label{tab:ea20-price-counterfactual-measures}\\
\toprule
Country & Price measures in Arregui et al. Annex II & Simulation status & Price counterfactuals simulated in the programs & Coverage notes \\
\midrule
\endfirsthead
\toprule
Country & Price measures in Arregui et al. Annex II & Simulation status & Price counterfactuals simulated in the programs & Coverage notes \\
\midrule
\endhead
\midrule
\multicolumn{5}{r}{\emph{Continued on next page}}\\
\endfoot
\bottomrule
\endlastfoot
Malta & Regulated/subsidized energy prices. & Not simulated in current programs &  &  \\
France & Gas/electricity tariff shields; fuel rebate. & Simulated and plotted & Electricity (CP0451): regulated-price / CRE existing replication for the French tariff shield [existing France replication].; Gas (CP0452): regulated-price / CRE existing replication for the French gas tariff freeze [existing France replication].; Transport fuels (CP0722): French pump fuel discount reconstruction, 18/30/10 cents per litre in 2022 [fuel rebate replication]. &  \\
Slovakia & Heat, electricity and gas price caps. & Identified in Bruegel/program audit; missing statutory schedule or external series &  & Pending statuses: needs statutory price schedule \\
Austria & Electricity price cap / block tariff; energy tax reduction. & Identified in Bruegel/program audit; missing statutory schedule or external series &  & Pending statuses: needs statutory price schedule \\
Greece & Electricity/gas unit subsidies; social tariffs; fuel support. & Simulated and plotted & Electricity (CP0451): unit schedule; July 2022 household electricity subsidy: EUR 200/MWh.; August 2022 household electricity subsidy: EUR 337/MWh.; September 2022 household electricity subsidy: EUR 639/MWh. [2022-07 to 2022-09; unit-subsidy rechained].; Gas (CP0452): unit schedule; July 2022 natural gas subsidy: EUR 30/MWh.; October 2022 natural gas subsidy: EUR 90/MWh. [2022-07 to 2022-10; unit-subsidy rechained]. &  \\
Lithuania & Electricity and gas compensation/subsidy schedules. & Simulated and plotted & Electricity (CP0451): maximum statutory unit compensation; Second half 2022: compensation cap of EUR 0.09/kWh above EUR 0.24/kWh.; First half 2023: compensation cap of EUR 0.285/kWh above EUR 0.28/kWh. [2022-07 to 2023-06; unit-subsidy rechained].; Gas (CP0452): statutory unit compensation converted from m3; Second half 2022: gas compensation of EUR 0.54/m3 converted with 10.55 kWh/m3.; First half 2023: gas compensation of EUR 0.99/m3 converted with 10.55 kWh/m3. [2022-07 to 2023-06; unit-subsidy rechained]. &  \\
Italy & Gas VAT cut; system-charge reductions; social energy bonus; fuel tax measures. & Simulated and plotted & Gas (CP0452): Gas VAT reduction from ordinary 22\% to 5\%; system-charge reductions/social bonuses not included. [2022-01 to 2023-12; VAT-exclusive rechained].; Electricity (CP0451): aggregate social-bonus equivalent; EUR 1.313bn / 3.766m bonuses / 2000 kWh, multiplied by an assumed 14.7\% eligible-household share. [2022-01 to 2022-12; unit-subsidy rechained].; Gas (CP0452): aggregate social-bonus equivalent; EUR 849m / 2.441m bonuses / 10550 kWh, multiplied by an assumed 9.5\% eligible-household share. [2022-01 to 2022-12; unit-subsidy rechained]. & Italy social-bonus unit equivalents are plotted tests, not final HICP counterfactuals. \\
Croatia & Gas/heat VAT cut; household electricity and heat caps/block support. & Simulated and plotted & Gas (CP0452): VAT reduction on gas from the standard 25\% to 5\%; heating treatment shown separately when HICP CP0455 is available. [2022-04 to 2023-03; VAT-exclusive rechained].; Heat energy (CP0455): VAT reduction on heating from the reduced 13\% rate to 5\%; depends on CP0455 availability. [2022-04 to 2023-03; VAT-exclusive rechained]. &  \\
Latvia & Electricity OIK/network support; energy compensation including gas/fuels. & Identified in Bruegel/program audit; missing statutory schedule or external series &  & Pending statuses: needs consumption denominator \\
Netherlands & Energy VAT cut; 2023 price cap/block-style support. & Simulated and plotted & Electricity (CP0451): VAT reduction on energy from 21\% to 9\%; 2023 price cap not included. [2022-07 to 2022-12; VAT-exclusive rechained].; Gas (CP0452): VAT reduction on energy from 21\% to 9\%; 2023 price cap not included. [2022-07 to 2022-12; VAT-exclusive rechained]. &  \\
Portugal & Electricity VAT cut; gas/electricity/fuel tax and cap measures. & Simulated and plotted & Electricity (CP0451): Electricity VAT reduction from 13\% to 6\%; threshold/consumption constraints not modelled. [2022-10 to 2023-12; VAT-exclusive rechained]. &  \\
Germany & Gas VAT cut; gas/electricity price brakes. & Simulated and plotted & Gas (CP0452): VAT reduction on gas from 19\% to 7\%; gas price brake not included. [2022-10 to 2023-12; VAT-exclusive rechained]. &  \\
Luxembourg & Electricity/gas/fuel price stabilization and VAT/tax measures. & Identified in Bruegel/program audit; missing statutory schedule or external series &  & Pending statuses: needs statutory price schedule \\
Estonia & Electricity universal service/block tariff; network-fee cuts; energy excise deferrals. & Identified in Bruegel/program audit; missing statutory schedule or external series &  & Pending statuses: needs statutory price schedule \\
Belgium & VAT/excise cuts; social energy tariffs. & Simulated and plotted & Electricity (CP0451): VAT reduction on electricity from 21\% to 6\%; social tariff not included. [2022-03 to 2023-03; VAT-exclusive rechained].; Gas (CP0452): VAT reduction on gas from 21\% to 6\%; social tariff not included. [2022-04 to 2023-03; VAT-exclusive rechained]. &  \\
Slovenia & Petrol/diesel/heating-oil price regulation. & Benchmark excluded &  & Pending statuses: needs statutory price schedule A Slovenia fuel-cap synthetic-neighbour benchmark exists but is marked excluded by user, so it is not counted as a final simulated measure. \\
Spain & Electricity VAT/tax cuts; gas VAT cut; fuel rebate; Iberian gas cap/electricity counterfactual. & Simulated and plotted & Electricity (CP0451): Electricity VAT kept/reduced at 10\% relative to the normal 21\% rate. [2021-09 to 2022-06; VAT-exclusive rechained].; Electricity (CP0451): Electricity VAT reduction to 5\%; modelled against the normal 21\% rate. [2022-07 to 2023-06; VAT-exclusive rechained].; Gas (CP0452): Natural gas and biomass VAT reduction to 5\%; biomass is not separately modelled. [2022-10 to 2023-12; VAT-exclusive rechained].; Electricity (CP0451): Spain retail electricity regression plus IMF-calibrated Iberian cap counterfactual [existing Spain replication].; Transport fuels (CP0722): Spain fuel rebate reconstruction, EUR 0.20/litre in Apr-Dec 2022 [fuel rebate replication]. &  \\
Ireland & Electricity and gas VAT cuts. & Simulated and plotted & Electricity (CP0451): VAT reduction on electricity from 13.5\% to 9\%. [2022-05 to 2023-09; VAT-exclusive rechained].; Gas (CP0452): VAT reduction on gas from 13.5\% to 9\%. [2022-05 to 2023-09; VAT-exclusive rechained]. &  \\
Cyprus & Electricity VAT cut and tariff subsidies; fuel/heating-fuel excise cut. & Simulated and plotted & Electricity (CP0451): Household electricity VAT reduction from 19\% to 9\%; vulnerable 5\% rate not separately modelled. [2021-11 to 2022-08; VAT-exclusive rechained]. &  \\
Finland & Electricity VAT cut; biofuel obligation/fuel-price measure. & Simulated and plotted & Electricity (CP0451): Temporary electricity VAT reduction from 24\% to 10\%. [2023-01 to 2023-04; VAT-exclusive rechained]. &  \\
\end{longtable}
\par\vspace{0.35em}\footnotesize\noindent Sources: Bruegel national energy-policy dataset \citep{sgaravatti2021national}, Arregui et al. \citep{arregui2022targeted}, and authors' calculations from the counterfactual simulation programs.
\end{landscape}
